Hybrid platforms and bargaining

### 2023-06-15 src/presentation/io_day.qmd
Minor improvements to presentation

# Microfounding the profit function

@@

::: {.fragment}
Logit-like demand:
$$
x_{Ti} = \frac{\exp\left( \frac{v_T - p_{Ti}}{\mu} \right)}{A}
$$
:::
::: {.fragment}
where
$$
A = \int_0^{N_F} \exp\left( \frac{v_F - p_{Fi}}{\mu} \right) \mathrm{d}i + \int_0^{N_P} \exp\left( \frac{v_P - p_{Pi}}{\mu} \right) \mathrm{d}i + 1
$$
:::

::: {.crossref .fragment}
Utility function: @sec-utility-function
:::

## Production
::: {.fragment}
 - Assume that the platform prices its products as if they were made by seperate, competitive sellers
   - Possible interpretation: profit-maximizing subsidiaries
   - More importantly: "best case scenario"
:::
::: {.fragment}
 - The optimal price is an additive markup rule:
 $$
 \begin{align*}
 p^*_{Ti} &= c_T + \mu \\
 \pi^{v*}_{Ti} &= \mu \frac{\exp \left( \frac{v_T - c_T - \mu}{\mu} \right)}{A}
 \end{align*}
 $$
:::

@@

::: {.incremental}
 - **Benchmark**<br>platform can commit to any lump-sum entry fee $F_F$
 - **Bargaining**<br>total profits are shared according to the Shapley-value-based rule
:::

@@

::: {.incremental}
 - Optimal entry fee is independent of $N_P$:
 - Size of the aggregate is independent* of $N_P$:
   $$ A^* = \sqrt{\frac{\mu V_F}{I_F}} $$
 - **Consumer surplus** is also **independent*** of $N_P$:
   $$ CS^* \propto \log(A^*) $$
:::

@@

::: {.fragment}
 - In this paper
   - Endogenenize hybridness ($N_P$)
   - Relax separate pricing assumption
   - Alternative benchmarks (e.g. exogeneous split)
:::
::: {.fragment}
 - Longer term
   - Endogeneize consumer entry (possibly with consumers also engaging in the negotiation)
   - Apply this bargaining rule to other, similar settings
:::

>>>>

# A concrete example

@@

::: {.fragment}
Logit-like demand:
$$
x_{Ti} = \frac{\exp\left( \frac{v_T - p_{Ti}}{\mu} \right)}{A}
$$
:::
::: {.fragment}
where
$$
\scriptsize
A = \int_0^{N_F} \exp\left( \frac{v_F - p_{Fi}}{\mu} \right) \mathrm{d}i + \int_0^{N_P} \exp\left( \frac{v_P - p_{Pi}}{\mu} \right) \mathrm{d}i + 1
$$
:::

::: {.crossref}
Utility function: @sec-utility-function
:::

## Production
::: {.fragment}
 - Assume that the platform prices its products as if they were made by seperate, competitive sellers
   - Possible interpretation: profit-maximizing subsidiaries
   - More importantly: "best case scenario"
:::
::: {.fragment}
 - The optimal price is an additive markup rule:
 $$
 p^*_{Ti} &= c_T + \mu
 $$
:::

@@

::: {.incremental}
 - **Bargaining**<br>total profits are shared according to the Shapley-value-based rule
 - **Benchmark**<br>platform can commit to any lump-sum entry fee $F_F$
:::

@@

::: {.incremental}
 - Optimal entry fee is independent of $N_P$
 - Size of the aggregate is independent* of $N_P$:
   $$ A^* = \sqrt{\frac{\mu V_F}{I_F}} = \text{constant} $$
 - **Consumer surplus** is also **independent*** of $N_P$:
   $$ CS^* \propto \log(A^*) = \text{constant}$$
:::

@@

::: {.fragment}
 - In this paper
   - Endogenenize "hybridness" ($N_P$)
   - Relax separate product pricing assumption
   - Alternative benchmarks (e.g. exogeneous split)
:::
::: {.fragment}
 - Longer term
   - Endogeneize consumer entry
   - Three-sided negotiation (platform, firms, consumers)
   - Test bargaining through lab experiment
:::
